---
title: "Peer Assessment 1 - Reproducable Research"
output:
  html_document:
    keep_md: yes
---

Assignement 1
https://github.com/coursera-jm/RepData_PeerAssessment1

## Loading and preprocessing the data

Read in the data.
```{r}
# Define some options for knitr
knitr::opts_chunk$set(tidy=FALSE, fig.path='figures/')
# cleanup
rm(list=ls())
#read the csv file
rfile <- read.csv('activity.csv')
dim(rfile)
head(rfile)
str(rfile)
```
## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
total_of_steps_per_day <- sum(rfile$steps, na.rm = TRUE)
total_steps_each_day <- aggregate(formula=steps~date, 
                                  data=rfile, 
                                  FUN=sum, 
                                  na.rm=TRUE)
```
2. Generating the Histogram by each day
```{r}
hist(total_steps_each_day$steps,
     col = "blue",
     main = "Histogram of total steps per day",
     xlab = "Number of steps per day")
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(total_steps_each_day$steps)
median(total_steps_each_day$steps) 
```
## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
five_min_avg <- aggregate(formula=steps~interval, 
                          data=rfile, 
                          FUN=mean, 
                          na.rm=TRUE)
plot(x=five_min_avg$interval,
     y=five_min_avg$steps,
     type="l",
     main="Average Steps Per 5-Minute Interval",
     xlab="Interval",
     ylab="Number of Steps")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
intv <-five_min_avg$interval[which.max(five_min_avg$steps)]
intv # maximum interval
```
## Imputing missing values
1. Calculate and report the total number of missing values in the dataset 
```{r}
missing <- sum(is.na(rfile))
missing
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
library(plyr)
#Create a dataframe that holds the mean steps for weekdays and weekends.
mean_imput <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
rfile_impute <- ddply(rfile, ~interval, transform, steps = mean_imput(steps))
```
The following data set contains the sample data aggregated by date.
```{r}
rfile_impute.date <- aggregate(rfile_impute[1],
                               by=rfile_impute[2],
                               FUN=sum,
                               na.rm=TRUE)
hist(rfile_impute.date$steps,
     breaks=20,
     col = "blue",
     main = "Histogram of Total Number of Steps Taken per Day (Imputed Data)",
     xlab = "Steps per Day")

mean(rfile_impute.date$steps)   # mean
median(rfile_impute.date$steps) # median
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
#Adding new variable for weekend/weekday
rfile_impute$day <- "weekday"
rfile_impute$day[weekdays(as.Date(rfile_impute$date), abb=T) %in% c("Sat","Sun")] <- "weekend"
table(rfile_impute$day)
library(lattice)
mean_miss_day <- aggregate(steps~interval + day, 
                           data=rfile_impute, 
                           FUN="mean")
xyplot(steps ~ interval | day, 
       data=mean_miss_day, 
       type="l", 
       grid=T, 
       layout=c(1,2), 
       ylab="Number of steps", 
       xlab="Interval")

```


